Ecological performance of detecting data fabrication

### 2017-07-24 submission/manuscript.Rmd
Update manuscript.Rmd

In order to stifle attempts at data fabrication, improved detection of fabricated data is considered to deter such harmful attempts. 
Although deterrence theory dates back to the middle of the 17th century [@leviathan], its implementation has not occurred equally across fabrication, falsification, and plagiarism. Basically, deterrence theory stipulates that with increased risk of detection, the utility of scientific misconduct (for this context) will decrease and therefore fewer people will engage in such behaviors. This principle of deterrence has been implemented with plagiarism scanners, a development that already started a long time ago [e.g., @10.1109/13.28038]. In the last decade, detecting image manipulation has become one of the few forms of detecting data fabrication . The Journal of Cell Biology and EMBO scan each submitted image for potential manipulation [@The_Journal_of_Cell_Biology2015-vh;@10.1038/546575a], which greatly increases the risk of detecting (blatant) image manipulation. More recently, algorithms have been developed to automate the scanning of images for (subtle) manipulations [@10.1007/s11948-016-9841-7]. These developments in detecting image manipulation have increased detection risk during the pre-publication and post-publication phase by improving detection mechanisms and increasing the understanding of how images might be manipulated. Moreover, their application also helps researchers systematically evaluate research articles to estimate the extent of the problem of image manipulation [4\% of all papers are estimated to contain manipulated images, @10.1128/mBio.00809-16].

Statistical methods provide one way to improve detection of data fabrication in empirical research. Humans are notoriously bad at understanding and estimating probabilities [e.g., @10.1126/science.185.4157.1124;@10.1037/h0031322], which could manifest itself in the fundamentally probabilistic data they try to fabricate. 
When data are fabricated, probabilistic principles are easily violated if forgotten at the univariate level, bivariate level, trivatiate level, or beyond [@Haldane1948-nm]. Based on this idea, statistical methods that investigate whether the reported data are feasible under the theoretically probabilistic processes can be used to detect potential data fabrication.

@@

Nonetheless, prior to applying statistical methods to flag potentially problematic results, investigating whether such methods function well enough in detecting problems is required for responsible application. 
We hardly know how researchers might go about fabricating data. Cases such as Stapel, Fuji, Smeesters, and Sanna provide some insights, but are highly pre-selected (i.e., those who got caught/confessed) and as such, systematically biased. Relatively extensive descriptions in rare and partial autobiogrophical accounts provide little insight into the actual data fabrication process, except for the setting where it might take place [e.g., late at night when no one is around; @stapel_book]. However, trustworthiness of these accounts can be called into question. Additionally, the performance of methods to detect data fabrication is highly dependent on the unknown prevalence of data fabrication and the power to actually to detect data fabrication. Given that we do not know how researchers might fabricate data, the diagnosticity of these methods cannot realistically be assessed.

>>>>

In order to stifle attempts at data fabrication, improved detection of fabricated data is considered to deter such harmful attempts. Deterrence theory stipulates that with increased risk of detection, the utility of scientific misconduct (for this context) will decrease and therefore fewer people will engage in such behaviors. This principle of deterrence has been implemented with plagiarism scanners, a development that already started a long time ago [e.g., @10.1109/13.28038]. In the last decade, detecting image manipulation has become one of the few forms of detecting data fabrication. The Journal of Cell Biology and EMBO scan each submitted image for potential manipulation [@The_Journal_of_Cell_Biology2015-vh;@10.1038/546575a], which greatly increases the risk of detecting (blatant) image manipulation. More recently, algorithms have been developed to automate the scanning of images for (subtle) manipulations [@10.1007/s11948-016-9841-7]. These developments in detecting image manipulation have increased detection risk during the pre-publication and post-publication phase by improving detection mechanisms and increasing the understanding of how images might be manipulated. Moreover, their application also helps researchers systematically evaluate research articles to estimate the extent of the problem of image manipulation [4\% of all papers are estimated to contain manipulated images, @10.1128/mBio.00809-16].

Statistical methods provide another way to improve detection of data fabrication in empirical research. Humans are notoriously bad at understanding and estimating probabilities [e.g., @10.1126/science.185.4157.1124;@10.1037/h0031322], which could manifest itself in the fundamentally probabilistic data they try to fabricate. When data are fabricated, probabilistic principles are easily violated if forgotten at the univariate level, bivariate level, trivatiate level, or beyond [@Haldane1948-nm]. Based on this idea, statistical methods that investigate whether the reported data are feasible under the theoretically probabilistic processes can be used to detect potential data fabrication.

@@

Nonetheless, prior to applying statistical methods to flag potentially problematic results, investigating whether such methods function well enough in detecting problems is required for responsible application. We hardly know how researchers might go about fabricating data. Cases such as Stapel, Fuji, Smeesters, and Sanna provide some insights, but are highly pre-selected (i.e., those who got caught/confessed) and as such, systematically biased. Relatively extensive descriptions in rare and partial autobiogrophical accounts provide little insight into the actual data fabrication process. Additionally, the performance of methods to detect data fabrication is highly dependent on the unknown prevalence of data fabrication and the power to actually to detect data fabrication. Given that we do not know how researchers might fabricate data, the diagnosticity of these methods cannot realistically be assessed.
